Scalar Items in Embedded Position: {A}n Experimental Revisit

### 2013-01-21 text/paper/paper3.tex
paper3 - general discussion fixed

%Several worries and objections are conceivable that might cast doubt
%on the previous interpretation of our experimental data.

In the following, we would like to address potential alternative
explanations for the present results. We will argue that besides
these interpretation variants, the main point raised
above remains valid: neither traditionalism, nor grammaticalism
is able to account for the present pattern of results.

%some of these that we consider most
%pressing. Eventually, however, we conclude that the main point raised
%above remains valid: neither traditionalism, nor grammaticalism
%explains the observations even under alternative interpretations of the
%data.

First, it could be questioned whether our design reliably 
indicates the preference relation in (\ref{bsp:observed-preferences}) 
among readings of our target sentences. Thus, the incremental verification task
(\ivt) might simply select for the weakest possible reading. This idea
can be grounded in the observation that the preferred \lc-readings
of our target-related controls entail the dispreferred
\ec-readings. It could thus be argued that the difference in responses
for preference-related controls are due to the entailment properties,
not the known preference relation over \lc- and \ec-readings. However, this
does not explain why the distribution of responses in the target
conditions are obviously not governed by logical strength. In fact, it
is a very noteworthy point that although entailment relations among
readings differ between \as- and \es-sentences, the observed response
pattern does not. (This is what is at the heart of the problem for a
grammaticalism that relies on the strongest meaning hypothesis.)
\dn{Petra: ich glaub, den Punkt verstehe ich nicht ganz - sollte man hier
auch die Prosodie erwähnen? Phrasierung wirkt sich ja nicht auf entailment
relations aus, sondern wird normalerweise dem Syntax-Prosodie-Interface
zugeschrieben}

But maybe, a similar kind of objection goes, it is not feasible to
actually compare results from the preference-related controls to
results from our target conditions, because the kind of ``ambiguity''
is of a different kind. In the former we have a syntactic
ambiguity, in the latter a small number of competing pragmatic
enrichments. The former is a case of perceived ambiguity (some of our
subjects reported this in a \emph{Nachbefragungsbogen}\dn{English?}), 
the latter arguably is
not \citep[c.f.][]{GeurtsPouscoulous2009:Embedded-Implic}. So perhaps
even if the distribution of responses is indicative of the known
preference patterns in preference-related controls, this would not
necessarily mean that the distribution of responses in target
conditions is indicative of preference relations.

There are two kinds of reply to this objection (and others like
it). The first kind of reply starts with a concession. It is true that
we cannot conclude with certainty that our data implies that
(\ref{bsp:observed-preferences}) is the actual preference relation
among candidate readings of our target sentences. But we do not have
to either. Maybe in fact, the observed distribution has nothing to do
with preferences over readings. It does not matter. The main point is
that neither traditionalism nor grammaticalism predicts the observed
distribution of responses. The second kind of reply is fatalistic but
stubborn. If we assume that the \ivt is incapable of indicating
preferences among readings for pragmatically unspecified, then it
should at least be able to indicate availability of readings, should
it not? And if we deny the observed response patterns the power to
indicate preference, what should the information about availability of
readings be that we may safely infer from it? It should be that local
readings are available and that global readings are not. But that is
again bad news for both traditionalism and grammaticalism. The former
cannot explain the substantial number of local responses, the latter
doesn't explain the absence of global responses.

\begin{itemize}
\item pictures overemphasize contrast some/all
\item subjects tricked into local responses because they are at the end
  \begin{itemize}
  \item would only show availability after all
  \end{itemize}
\item typicality after all
\end{itemize}

\subsubsection{Bits and Pieces}
\label{sec:bits-pieces}

\begin{itemize}
\item The distribution of
responses found in previous studies could in part be explained by
differences in typicality of the pictures with respect to the possible
readings (see van Tiel). In the present case this kind of reasoning
would have to be extended to the typicalilty of parts of pictures. On
intuitive grounds, we take this kind of reasoning to be
implausible. We will now shortly discuss how graphical properties
might still affect judgments in the worst case. In the first step of
the \as-condition local and global readings cannot be judged,
yet. Therefore, graphical properties should not affect judgments if
participants have one of these readings in mind. If participants have
a literal reading in mind, the worst case scenario is that they
refrain from judging the sentence as true at this point due to some,
eg. atypical, graphical property of the picture. The situation is
similar at the second step. The local reading is still not
decidable. Therefore, graphical properties should not play a role if
participants have a local reading in mind. A positive judgment might,
however, be a delayed judgment of a literal reading. Again, the worst
case scenario for our purpose is that participants refrain from
answering because of graphical properties. At the last step the local
reading requires a negative judgment. Now, we could obtain delayed
negative responses from literal or global readings because the graph
as a whole is, e.g., logically true but atypical with respect to these
readings. These kinds of judgments would be indistinguishable from
local responses. Note, however, that we have no reason to believe that
typicality has large effects, because, given the results from van
Tiel, the picture as a whole is expected to be one of the more typical
situations for \as-sentences. Still, we cannot exclude the possibility
that graphical properties of the pictures might have the described
effect. In the \es-conditions this type of effect is, however,
entirely impossible since the literal and global reading is plainly
false given the initial parts as well as the whole graph. Summing up,
in the worst case, which we consider implausible, graphical effects
might lead to an enhanced number of local responses in the \as- as
compared to the \es-condition. We have to keep this possibility in
mind. We want to stress, however, that the design presented here
improves on previous studies with respect to possible graphical
effects.
\end{itemize}

@@

\begin{itemize}
\item future research: isolate consistent localists to probe into \emph{processing}
\end{itemize}

>>>>

There are a number of conceivable objections to our interpretation of
the presented results. We will consider the most pressing ones in the
following and conclude that the main point raised above remains valid:
neither traditionalism, nor grammaticalism is able to account for the
present pattern of results.

To begin with, it could be questioned whether our design reliably
indicates the preference relation in (\ref{bsp:observed-preferences})
among readings of our target sentences. After all, our incremental
verification task might simply select for the weakest possible
reading. This idea can be grounded in the observation that the
preferred \lc-readings of our target-related controls entail the
dispreferred \ec-readings. It could thus be argued that the difference
in responses for preference-related controls are due to the entailment
properties, not the known preference relation over \lc- and
\ec-readings. However, this does not explain why the distribution of
responses in the target conditions are obviously not governed by
logical strength. In fact, it is a very noteworthy point that although
entailment relations among readings differ between \as- and
\es-sentences, the observed response pattern does not. (This is what
is at the heart of the problem for a grammaticalism that relies on the
strongest meaning hypothesis.)

A related objection would be that maybe it is not feasible to actually
compare results from the preference-related controls to results from
our target conditions, because the source of different readings is
fundamentally different. In the former we have a syntactic ambiguity,
in the latter a small number of competing putative pragmatic
enrichments. The former is a case of perceived ambiguity (some of our
subjects reported this), the latter arguably is not
\citep[c.f.][]{GeurtsPouscoulous2009:Embedded-Implic}. So perhaps even
if the distribution of responses is indicative of the known preference
patterns in preference-related controls, this would not necessarily
mean that the distribution of responses in target conditions is
indicative of preference relations.

There are two kinds of reply to this objection (and others like
it). Firstly, it is true that we cannot conclude with certainty that
our data implies that (\ref{bsp:observed-preferences}) is the actual
preference relation among candidate readings of our target
sentences. But that does not matter. Even if the observed distribution
has nothing to do with preferences over readings, the main point
remains that neither traditionalism nor grammaticalism predicts the
observed distribution of responses. Secondly, even if we assume that
the incremental verification task is incapable of indicating
preferences among pragmatic readings of the relevant kind, then it
should at least be able to indicate availability of readings. So if
we deny the observed response patterns the power to indicate
preference, what should the information about availability of readings
be that we may safely infer from it? It should be that local readings
are available and that global readings are not. But that is again not
predicted by either traditionalism or grammaticalism. The former cannot
explain the substantial number of local responses, the latter doesn't
explain the absence of global responses.

A third objection could be that pictures in our tasks may unduly
overemphasize the contrast between cases where items are connected
with all or merely some but not all of their surrounding shapes. This
is essentially the line of argument pursued by
\citet{GeurtsTielvan-Tiel2013:Scalar-expressi} against the data
reported by \citet{ChemlaSpector2010:Experimental-Ev} in support of
local readings of
\es-sentences. \citeauthor{GeurtsTielvan-Tiel2013:Scalar-expressi}
show that the strength of agreement to an \es-sentence on a 7-point
Likert-scale depends on a visual contrast between, on the one hand,
one element connected to some but not all of its surrounding shapes,
and, on the other hand, a variable number of elements connected to all
of their surrounding shapes. When two universally connected elements
are present, the mean rate of agreement is higher than when only one
is present. So, maybe our displays similarly provoked local readings
because of an overemphasized visual contrast.

This cannot be a plausible explanation for the number of local
responses we found in our study. Although it is true that the critical
position for judging the local reading is a picture with a high visual
contrast in the above sense (see Figure~\ref{fig:exseqES}), a purely
contrast-based explanation of the local responses in our data is not
feasible for two reasons. The first is that visual contrast does not
explain how subjects ever even got past the second critical position
where the sentence can be judged false under its literal reading. At
this position there is a visual contrast of the alleged kind but it is
as small as possible. So something must have motivated subjects
already at this position to continue unraveling, given that they were
generally able of exiting at the earliest possible location. A second
argument that casts doubt on a contrast-based explanation is that our
study elicited categorical truth-value judgements. Unlike in
\citeauthor{ChemlaSpector2010:Experimental-Ev}'s
(\citeyear{ChemlaSpector2010:Experimental-Ev}) and
\citeauthor{GeurtsTielvan-Tiel2013:Scalar-expressi}'s
(\citeyear{GeurtsTielvan-Tiel2013:Scalar-expressi}) experiments, we
did not record degrees of agreement with a statement. But it is hardly
plausible that the visual contrast in a picture like in
\ref{fig:subfig1} alone is enough to overrule a truth-value judgment
so as to contradict the semantic meaning (even when there is no
intonational markedness to support such a
reinterpretation). Ultimately, however, even if the local answers in
our study are contrast effects, the verdict remains that neither
traditionalism can not account for that piece of the data unless aided
by yet another auxiliary and grammaticalism cannot explain why it is
relatively dispreferred.

A fourth major conceivable objection maintains that we cannot rule out
that the way we realized the intonational markedness in the accented
conditions was not the right kind of intonational
markedness. Therefore the intonational markedness hypothesis that
links an increase in local readings to intonation might not be
falsified by our results. Although true, this still does not help
vindicate traditionalism, for even without marked intonation local
readings were available.

Finally, we should address the obvious worry that maybe our results
could have been the product of typicality after all. Remember that
\citet{Tielvan-Tiel2012:Embedded-Scalar} and
\citet{GeurtsTielvan-Tiel2013:Scalar-expressi} argue that the
distribution of responses to \as-sentences found in previous studies
could in part be explained by differences in typicality of the
pictures with respect to the possible readings. In the present case
this kind of reasoning would have to be extended to either cover the
typicality of parts of pictures or to figure in subjects' expectations
of pictorial typicality given their uncertainty. On intuitive grounds,
we take either extension to be far-fetched and implausible. But even
if this line of explanation can be plausibly extended to incrementally
revealed picture verification, this does not as such save the day for
traditionalism or grammaticalism, for the same reasons as given above.

% Still, graphical properties might affect judgments in the worst
% case. In the first step of the \as-condition local and global readings
% cannot be judged, yet. Therefore, graphical properties should not
% affect judgments if participants have one of these readings in
% mind. If participants have a literal reading in mind, the worst case
% scenario is that they refrain from judging the sentence as true at
% this point due to some, say, atypical, graphical property of the
% picture. The situation is similar at the second step. The local
% reading is still not decidable. Therefore, graphical properties should
% not play a role if participants have a local reading in mind. A
% positive judgment might, however, be a delayed judgment of a literal
% reading. Again, the worst case scenario for our purpose is that
% participants refrain from answering because of graphical
% properties. At the last step the local reading requires a negative
% judgment. Now, we could obtain delayed negative responses from literal
% or global readings because the graph as a whole is, e.g., logically
% true but atypical with respect to these readings. These kinds of
% judgments would be indistinguishable from local responses. Note,
% however, that we have no reason to believe that typicality has large
% effects, because, given the results from van Tiel, the picture as a
% whole is expected to be one of the more typical situations for
% \as-sentences. Still, we cannot exclude the possibility that graphical
% properties of the pictures might have the described effect. In the
% \es-conditions this type of effect is, however, entirely impossible
% since the literal and global reading is plainly false given the
% initial parts as well as the whole graph. Summing up, in the worst
% case, which we consider implausible, graphical effects might lead to
% an enhanced number of local responses in the \as- as compared to the
% \es-condition. We have to keep this possibility in mind. We want to
% stress, however, that the design presented here improves on previous
% studies with respect to possible graphical effects.

@@

\begin{itemize}
\item future research: isolate consistent localists to probe into \emph{processing}
\item production study
\end{itemize}
